### 2021-07-20 ch/7-discussion.tex
Small improvements

% ======== giving my social perspective viewpoint ========
As AI models (as opposed to mere \ac{ML} models) become more powerful,
the biases as presented in this thesis will likely become less relevant.
However, we can argue that generalisations of these biases will still be applicable.
Spurious correlations in the training data might become less relevant,
because a powerful AI can do its own physics experiments
and thus will be able to figure out the causal structure of our world on its own.
However, what is still relevant is the existence of robust, analysable world models,
which it is possible to reason on and do planning with.
The potential trap here is not that the world model is \emph{wrong},
but that it fails to have a useful structure for other tasks.
Furthermore, learning about the physical world will not in and of itself be enough to
become aligned with \emph{human values},
which is here a catch-all term for
what morality is or what good is and what wrong is and what good things are,
the things that we want more of, and what bad things are.
Such questions may appear very different from the issue of label bias
and objectives like demographic parity,
but if we take the problem of classical algorithmic fairness to be
that we want the model to treat our fellow humans right,
then this is just the generalisation of that.
% ==============

>>>>

% ======== giving my social perspective viewpoint ========
As AI models (as opposed to mere \ac{ML} models) become more powerful,
the biases as presented in this thesis will likely become less relevant.
However, we can argue that more abstract forms of these problems will still be present.
Spurious correlations in the training data might become less relevant --
because a powerful AI can do its own physics experiments
and thus will be able to figure out the causal structure of our world on its own --
but what is still relevant is the existence of robust, analysable world models,
which it is possible to reason on and do planning with.
The potential trap here is not that the world model is \emph{wrong} per se,
but that it fails to have a useful structure for other tasks.
Furthermore, learning about the physical world will not in and of itself be enough to
become aligned with \emph{human values},
which is here a catch-all term for
what morality is or what good is and what wrong is.
% the things that we want more of, and what bad things are.
Such questions may appear very different from the issue of label bias
and objectives like demographic parity,
but if we take the problem of classical algorithmic fairness to be
that we want the model to treat our fellow humans right,
then this is just the generalisation of that.
% ==============

### 2021-07-20 paper1/paper1.tex
Small improvements

>>>>

Furthermore, as currently formulated, the approach is only compatible with likelihood-based models
(which covers neural networks),
but the only real requirement is reasonably-well calibrated probabilities from the used model,
which are needed to compute the expected target labels.
Thus, future work could extend it to other kinds of models, like \acp{SVM},
for which solutions exists to transform the output to a probability;
the most common being Platt scaling \citep{platt1999probabilistic},
which applies a logistic transformation with two free parameters to the output of the \ac{SVM}.
The free parameters have to be estimated by iterating over the whole training set.
The remaining (substantial) challenge is then
to integrate the loss computed from these probabilities into the training of the \ac{SVM}.

### 2021-07-20 paper3/paper3.tex
Small improvements

% \textbf{Disentangled representations via adversarial autoencoders}
%We want to learn a disentanglement using a perfect dataset as the supervision signal.
%
%To achieve this, our semi-supervised disentanglement learning framework (fig. \ref{fig:architecture} ) maps data points into independent factors of variations such that our training dataset $D_{\text{tr}}$ and our deployment set $D_c$ are indistinguishable; i.e., we would like transform $D_{\text{tr}}$ such that the independence condition $y\perp s$ holds.
%
%However, the deployment set itself $D_c$ might not amount to a suitable target distribution in which the independence condition holds.
%
%The idea is to make the deployment set $D_c$ approximately perfect by balancing it according to the cluster IDs, forming the set $D_p$.
%
% We create disentangled representations by training four separate neural networks, which we denote $f$, $g$, $h$ and $l$.
% %
% We use mean squared error for the reconstruction loss, and the binary cross entropy error for the discriminator loss.
% %
% For a dataset with mixed-typed attributes, we use a combination of mean squared error and (binary) cross entropy error for the reconstruction loss.
% %
% We optimise the proposed model using a scheduled update scheme where we freeze the weights of the autoencoder and predictor modules when we update the weights of the discriminator, and vice versa.
%
\section{Conclusion}
We have highlighted the problem that systematic bias can result in one or more subgroups having zero labelled data, and by doing so hope to have stimulated serious consideration for it (even if to be dismissed) when planning, building, evaluating and regulating machine learning systems.
We propose a two-step approach for addressing the resulting zero-label stratification problem.
First, we construct perfect bags from an unlabelled deployment set via  clustering. 
Second, we learn a disentangled representations using the perfect bags for adversarial distribution-matching. 
We empirically validate our framework on the Coloured MNIST, CelebA and Adult Income datasets, and find evidence that it is possible to maintain high performance on the subgroups with zero training data. 
We analyse our approach using the disentanglement calculus of \citet{ShuCheKumErmetal20} that relies on the notions of restrictiveness and consistency, and show that we can derive some guarantees.
Future work includes further restricting the amount of information that can be encoded in the bias factor, and the limitations that we mentioned in sec.~\ref{sec:limitations}.
Furthermore, the literature on clustering can potentially provide further insights on theoretical guarantees for clustering. This was not a focus in this paper as the disentangling aspect is more crucial to the approach.

>>>>

% \textbf{Disentangled representations via adversarial autoencoders}
%We want to learn a disentanglement using a perfect dataset as the supervision signal.
%
%To achieve this, our semi-supervised disentanglement learning framework (fig. \ref{fig:architecture} ) maps data points into independent factors of variations such that our training dataset $D_{\text{tr}}$ and our deployment set $D_c$ are indistinguishable; i.e., we would like transform $D_{\text{tr}}$ such that the independence condition $y\perp s$ holds.
%
%However, the deployment set itself $D_c$ might not amount to a suitable target distribution in which the independence condition holds.
%
%The idea is to make the deployment set $D_c$ approximately perfect by balancing it according to the cluster IDs, forming the set $D_p$.
%
% We create disentangled representations by training four separate neural networks, which we denote $f$, $g$, $h$ and $l$.
% %
% We use mean squared error for the reconstruction loss, and the binary cross entropy error for the discriminator loss.
% %
% For a dataset with mixed-typed attributes, we use a combination of mean squared error and (binary) cross entropy error for the reconstruction loss.
% %
% We optimise the proposed model using a scheduled update scheme where we freeze the weights of the autoencoder and predictor modules when we update the weights of the discriminator, and vice versa.
%
\section{Conclusion}
We have highlighted the problem that systematic bias can result in one or more subgroups having zero labelled data, and by doing so hope to have stimulated serious consideration for it (even if to be dismissed) when planning, building, evaluating and regulating machine learning systems.
We propose a two-step approach for addressing the resulting zero-label stratification problem.
First, we construct perfect bags from an unlabelled deployment set via  clustering. 
Second, we learn a disentangled representations using the perfect bags for adversarial distribution-matching. 
We empirically validate our framework on the Coloured MNIST, CelebA and Adult Income datasets, and find evidence that it is possible to maintain high performance on the subgroups with zero training data. 
We analyse our approach using the disentanglement calculus of \citet{ShuCheKumErmetal20} that relies on the notions of restrictiveness and consistency, and show that we can derive some guarantees.
We presented our approach in the context of biased data,
but it could be used as a general-purpose method for learning invariances
Future work includes further restricting the amount of information that can be encoded in the bias factor, and the limitations that we mentioned in sec.~\ref{sec:limitations}.
Furthermore, the literature on clustering can potentially provide further insights on theoretical guarantees for clustering. This was not a focus in this paper as the disentangling aspect is more crucial to the approach.
